Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2021-04-12 analysis/eval-performance.Rmd

flextable::flextable(df_tab)
```

@@

```{r eval-performance-1, warning=FALSE}
table1 <- df_perf %>%
  group_by(learner.id, task.id, filter) %>%
  slice(which.min(rmse_aggr)) %>%
  mutate(rsq = round(rsq, 3)) %>%
  dplyr::rename(
    "Model" = learner_group,
    "Learner ID" = learner.id,
    "Task" = task.id,
    "Filter" = filter,
    "RMSE" = rmse_aggr,
    "SE (RMSE)" = se,
    "R^2" = rsq,
  ) %>%
  ungroup() %>%
  mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
  select(-iter, -`Learner ID`, -expvar) %>%
  relocate(RMSE, .after = Filter) %>%
  relocate(`SE (RMSE)`, .after = RMSE) %>%
  mutate(RMSE = round(RMSE, 3)) %>%
  arrange(RMSE)

@@

table1 %>%
  flextable() %>%
  autofit()
```

@@

```{r eval-performance-2, warning=FALSE}
table2 <- df_perf %>%
  group_by(learner_group) %>%
  slice(which.min(rmse_aggr)) %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  arrange(rmse_aggr) %>%
  mutate(rsq = round(rsq, 3)) %>%
  dplyr::rename(
    "Model" = learner_group,
    "Learner ID" = learner.id,
    "Task" = task.id,
    "Filter" = filter,
    "RMSE" = rmse_aggr,
    "SE (RMSE)" = se,
    "R^2" = rsq,
  ) %>%
  select(-`Learner ID`) %>%
  relocate(RMSE, .after = Filter) %>%
  relocate(`SE (RMSE)`, .after = RMSE) %>%
  select(-iter, -expvar)

@@

table2 %>%
  flextable() %>%
  autofit()
```

@@

```{r eval-performance-3, warning=FALSE}
table3 <- df_perf %>%
  group_by(learner.id, task.id, filter) %>%
  slice(which.min(rmse_aggr)) %>%
  mutate(rsq = round(rsq, 3)) %>%
  dplyr::rename(
    "Model" = learner_group,
    "Learner ID" = learner.id,
    "Task" = task.id,
    "Filter" = filter,
    "RMSE" = rmse_aggr,
    "R^2" = rsq,
    "SE" = se,
    "SE (RMSE)" = se
  ) %>%
  ungroup() %>%
  mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
  select(-iter, -`Learner ID`, -expvar) %>%
  relocate(RMSE, .after = Filter) %>%
  relocate(`SE (RMSE)`, .after = RMSE) %>%
  mutate(RMSE = round(RMSE, 3)) %>%
  arrange(RMSE)

@@

table3 %>%
  ungroup() %>%
  arrange(desc(RMSE)) %>%
  slice(1:10) %>%
  flextable() %>%
  autofit()
```

@@

```{r filter-effect-all-vs-borda-filter, warning=FALSE, dev = c("png", "pdf")}
results_aggr2 <- df_perf %>%
  na.omit() %>%
  filter(learner_group != "Ridge-MBO") %>%
  filter(learner_group != "Lasso-MBO") %>%
  mutate(learner_group = recode_factor(learner_group, `XGBoost` = "XG")) %>%
  group_by(learner_group, task.id, filter) %>%
  # we actually took the mean already in chunk 'aggr-perf'. This is only to get
  # summarise() working
  summarise(perf = mean(rmse_aggr)) %>%
  ungroup() %>%
  # we need to reverse the order on porpuse here so that ggplot reverses it
  # again later
  mutate(learner_group = fct_relevel(learner_group, "XGBoost", "SVM", "RF"))

@@

df_tbl <- getBMRTuneResults(bmr_inspect_tune, as.df = TRUE) %>%
  dplyr::rename(Learner = learner.id) %>%
  dplyr::rename(Plot = iter) %>%
  dplyr::rename(RMSE = rmse.test.mean) %>%
  dplyr::rename("Features (\\%)" = fw.perc) %>%
  dplyr::mutate(`Test Plot` = as.character(Plot)) %>%
  dplyr::mutate(`Test Plot` = forcats::fct_recode(`Test Plot`,
    Oiartzun = "4", Luiando = "3",
    Laukiz1 = "1", Laukiz2 = "2"
  )) %>%
  dplyr::mutate(Learner = forcats::fct_recode(Learner,
    "RF \\\\ Car" = "RF MBO Car", "XGB \\\\ Borda" = "XGBOOST MBO Borda",
    "SVM \\\\ Relief" = "SVM MBO Relief"
  )) %>%
  dplyr::mutate(Learner = as.character(Learner)) %>%
  dplyr::mutate(`Test Plot` = as.character(`Test Plot`)) %>%
  dplyr::mutate("\\#" = case_when(
    `Test Plot` == "Laukiz1" ~ ceiling(`Features (\\%)` * 479),
    `Test Plot` == "Laukiz2" ~ ceiling(`Features (\\%)` * 451),
    `Test Plot` == "Luiando" ~ ceiling(`Features (\\%)` * 300),
    `Test Plot` == "Oiartzun" ~ ceiling(`Features (\\%)` * 529),
  )) %>%
  dplyr::group_by(Learner) %>%
  dplyr::select(Learner, `Test Plot`, "Features (\\%)", "\\#", RMSE)

@@

table4 <- df_tbl %>%
  xtable::xtable(
    type = "latex",
    caption = "Selected feature portions during tuning for selected learner-filter settings across folds for task HR-NRI-VI, sorted ascending by RMSE",
    label = "tab:tune-perc-sel-features",
    digits = c(0, 0, 0, 5, 0, 2)
  )

@@

table4 %>%
  flextable() %>%
  autofit() %>%
  fontsize(size = 13)
```

@@

```{r}
getBMRTuneResults(bmr_inspect_tune, as.df = TRUE) %>%
  dplyr::rename(Learner = learner.id) %>%
  dplyr::rename(Plot = iter) %>%
  dplyr::rename(RMSE = rmse.test.mean) %>%
  dplyr::rename("Features (%)" = fw.perc) %>%
  dplyr::mutate(Plot = as.character(Plot)) %>%
  dplyr::mutate(Plot = forcats::fct_recode(Plot,
    Oiartzun = "2", Luiando = "3",
    Laukiz1 = "1", Laukiz2 = "4"
  )) %>%
  dplyr::mutate(Learner = forcats::fct_recode(Learner,
    "RF Car" = "RF MBO Car", "XGBoost Borda" = "XGBOOST MBO Borda",
    "SVM Relief" = "SVM MBO Relief"
  )) %>%
  dplyr::mutate(Learner = as.character(Learner)) %>%
  dplyr::mutate(Plot = as.character(Plot)) %>%
  dplyr::group_by(Learner) %>%
  dplyr::summarise(
    "Mean (Features (%))" = mean(`Features (%)`),
    "SD (Features (%))" = sd(`Features (%)`)
  ) %>%
  flextable() %>%
  autofit() %>%
  fontsize(size = 13)
```

>>>>

DT::datatable(df_tab)
```

@@

```{r eval-performance-1, warning=FALSE}
table1 <- df_perf %>%
  group_by(learner.id, task.id, filter) %>%
  slice(which.min(rmse_aggr)) %>%
  mutate(rsq = round(rsq, 3)) %>%
  dplyr::rename(
    "Model" = learner_group,
    "Learner ID" = learner.id,
    "Task" = task.id,
    "Filter" = filter,
    "RMSE" = rmse_aggr,
    "SE (RMSE)" = se,
    "$R^2$" = rsq,
  ) %>%
  ungroup() %>%
  mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
  select(-iter, -`Learner ID`, -expvar) %>%
  relocate(RMSE, .after = Filter) %>%
  relocate(`SE (RMSE)`, .after = RMSE) %>%
  mutate(RMSE = round(RMSE, 3)) %>%
  arrange(RMSE)

@@

DT::datatable(table1)
```

@@

```{r eval-performance-2, warning=FALSE}
table2 <- df_perf %>%
  group_by(learner_group) %>%
  slice(which.min(rmse_aggr)) %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  arrange(rmse_aggr) %>%
  mutate(rsq = round(rsq, 3)) %>%
  dplyr::rename(
    "Model" = learner_group,
    "Learner ID" = learner.id,
    "Task" = task.id,
    "Filter" = filter,
    "RMSE" = rmse_aggr,
    "SE (RMSE)" = se,
    "$R^2$" = rsq,
  ) %>%
  select(-`Learner ID`) %>%
  relocate(RMSE, .after = Filter) %>%
  relocate(`SE (RMSE)`, .after = RMSE) %>%
  select(-iter, -expvar)

@@

DT::datatable(table2)
```

@@

```{r eval-performance-3, warning=FALSE}
table3 <- df_perf %>%
  group_by(learner.id, task.id, filter) %>%
  slice(which.min(rmse_aggr)) %>%
  mutate(rsq = round(rsq, 3)) %>%
  dplyr::rename(
    "Model" = learner_group,
    "Learner ID" = learner.id,
    "Task" = task.id,
    "Filter" = filter,
    "RMSE" = rmse_aggr,
    "$R^2$" = rsq,
    "SE" = se,
    "SE (RMSE)" = se
  ) %>%
  ungroup() %>%
  mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
  select(-iter, -`Learner ID`, -expvar) %>%
  relocate(RMSE, .after = Filter) %>%
  relocate(`SE (RMSE)`, .after = RMSE) %>%
  mutate(RMSE = round(RMSE, 3)) %>%
  arrange(RMSE)

@@

DT::datatable(table3)
```

@@

```{r filter-effect-all-vs-borda-filter, warning=FALSE, dev = c("png", "pdf"), out.width = "70%"}
results_aggr2 <- df_perf %>%
  na.omit() %>%
  filter(learner_group != "Ridge-MBO") %>%
  filter(learner_group != "Lasso-MBO") %>%
  mutate(learner_group = recode_factor(learner_group, `XGBoost` = "XG")) %>%
  group_by(learner_group, task.id, filter) %>%
  # we actually took the mean already in chunk 'aggr-perf'. This is only to get
  # summarise() working
  summarise(perf = mean(rmse_aggr)) %>%
  ungroup() %>%
  # we need to reverse the order on porpuse here so that ggplot reverses it
  # again later
  mutate(learner_group = fct_relevel(learner_group, "XGBoost", "SVM", "RF"))

@@

df_tbl <- getBMRTuneResults(bmr_inspect_tune, as.df = TRUE) %>%
  dplyr::rename(Learner = learner.id) %>%
  dplyr::rename(Plot = iter) %>%
  dplyr::rename(RMSE = rmse.test.mean) %>%
  dplyr::rename("Features (\\%)" = fw.perc) %>%
  dplyr::mutate(`Test Plot` = as.character(Plot)) %>%
  dplyr::mutate(`Test Plot` = forcats::fct_recode(`Test Plot`,
    Oiartzun = "4", Luiando = "3",
    Laukiz1 = "1", Laukiz2 = "2"
  )) %>%
  dplyr::mutate(Learner = forcats::fct_recode(Learner,
    "RF \\\\ Relief" = "RF MBO Relief", "XGB \\\\ CMIM" = "XGBOOST MBO CMIM",
    "SVM \\\\ Relief" = "SVM MBO Relief"
  )) %>%
  dplyr::mutate(Learner = as.character(Learner)) %>%
  dplyr::mutate(`Test Plot` = as.character(`Test Plot`)) %>%
  dplyr::mutate("\\#" = case_when(
    `Test Plot` == "Laukiz1" ~ ceiling(`Features (\\%)` * 559),
    `Test Plot` == "Laukiz2" ~ ceiling(`Features (\\%)` * 451),
    `Test Plot` == "Luiando" ~ ceiling(`Features (\\%)` * 301),
    `Test Plot` == "Oiartzun" ~ ceiling(`Features (\\%)` * 497),
  )) %>%
  dplyr::group_by(Learner) %>%
  dplyr::select(Learner, `Test Plot`, "Features (\\%)", "\\#", RMSE)

@@

table4 <- df_tbl %>%
  xtable::xtable(
    type = "latex",
    caption = "Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted ascending by RMSE.",
    label = "tab:tune-perc-sel-features",
    digits = c(0, 0, 0, 5, 0, 2)
  )

@@

DT::datatable(table4)
```

@@

```{r}
getBMRTuneResults(bmr_inspect_tune, as.df = TRUE) %>%
  dplyr::rename(Learner = learner.id) %>%
  dplyr::rename(Plot = iter) %>%
  dplyr::rename(RMSE = rmse.test.mean) %>%
  dplyr::rename("Features (%)" = fw.perc) %>%
  dplyr::mutate(Plot = as.character(Plot)) %>%
  dplyr::mutate(Plot = forcats::fct_recode(Plot,
    Oiartzun = "2", Luiando = "3",
    Laukiz1 = "1", Laukiz2 = "4"
  )) %>%
  dplyr::mutate(Learner = forcats::fct_recode(Learner,
    "RF Car" = "RF MBO Car", "XGBoost Borda" = "XGBOOST MBO Borda",
    "SVM Relief" = "SVM MBO Relief"
  )) %>%
  dplyr::mutate(Learner = as.character(Learner)) %>%
  dplyr::mutate(Plot = as.character(Plot)) %>%
  dplyr::group_by(Learner) %>%
  dplyr::summarise(
    "Mean (Features (%))" = mean(`Features (%)`),
    "SD (Features (%))" = sd(`Features (%)`)
  ) %>%
  DT::datatable()
```

### 2021-04-12 analysis/feature-importance.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 8.5,
  fig.asp = 0.66,
  out.width = "100%",
  echo = FALSE
)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 8.5,
  fig.asp = 0.66,
  out.width = "70%",
  echo = FALSE
)

### 2021-04-12 docs/00-manuscripts/ieee/tune-perc-sel-features.tex

% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{table}[ht!]
\centering
\caption{Selected feature portions during tuning for selected learner-filter settings across folds for task HR-NRI-VI, sorted ascending by RMSE} 
\label{tab:tune-perc-sel-features}
\begin{tabular}{llrrr}
  \toprule
Learner & Test Plot & Features (\%) & \# & RMSE \\ 
  \midrule
\multirow{4}{*}{\specialcell{RF \\ Car}} & Laukiz1 & 0.14212 & 69 & 41.48 \\ 
   & Laukiz2 & 0.21934 & 99 & 16.47 \\ 
   & Luiando & 0.02878 & 9 & 46.33 \\ 
   & Oiartzun & 0.00279 & 2 & 31.82 \\ 
  \midrule\multirow{4}{*}{\specialcell{XGB \\ Borda}} & Laukiz1 & 0.01502 & 8 & 15.48 \\ 
   & Laukiz2 & 0.66619 & 301 & 44.39 \\ 
   & Luiando & 0.10428 & 32 & 40.86 \\ 
   & Oiartzun & 0.66651 & 353 & 34.15 \\ 
  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.00062 & 1 & 35.49 \\ 
   & Laukiz2 & 0.04106 & 19 & 31.03 \\ 
   & Luiando & 0.20735 & 63 & 37.19 \\ 
   & Oiartzun & 0.23928 & 127 & 14.89 \\ 
   \bottomrule
\end{tabular}
\end{table}

>>>>

% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{table}[ht!]
\centering
\caption{Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted ascending by RMSE.} 
\label{tab:tune-perc-sel-features}
\begin{tabular}{llrrr}
  \toprule
Learner & Test Plot & Features (\%) & \# & RMSE \\ 
  \midrule
\multirow{4}{*}{\specialcell{RF \\ Relief}} & Laukiz1 & 0.99973 & 559 & 41.73 \\ 
   & Laukiz2 & 0.93371 & 422 & 17.12 \\ 
   & Luiando & 0.00216 & 1 & 33.21 \\ 
   & Oiartzun & 0.00004 & 1 & 37.20 \\ 
  \midrule\multirow{4}{*}{\specialcell{XGB \\ CMIM}} & Laukiz1 & 0.30849 & 173 & 33.36 \\ 
   & Laukiz2 & 0.98122 & 443 & 45.59 \\ 
   & Luiando & 0.83613 & 252 & 41.01 \\ 
   & Oiartzun & 0.00515 & 3 & 15.24 \\ 
  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.00062 & 1 & 35.49 \\ 
   & Laukiz2 & 0.04106 & 19 & 31.03 \\ 
   & Luiando & 0.20735 & 63 & 37.19 \\ 
   & Oiartzun & 0.23928 & 119 & 14.89 \\ 
   \bottomrule
\end{tabular}
\end{table}
